Universal Anonymous Signatures

### 2022-10-03 prism-backend/research/gsac/introduction-paper.tex
Added back issuance unforgeability to fix issuance flaw. Updated rest of the paper. Proofs can be improved.
Pass.

%% Previous works summary w/ refs directly related contribs of UAS
%% (or, tech approaches of main precedents)
\paragraph{Evolution of AC and GS.} %
AC and GS systems have followed separate paths, with some exceptions.
Functionality-wise, Attribute-Based Group Signatures add attributes to the
credentials issued to group members, who can later use them to prove pre-fixed
claims \cite{emo09,aa14}. Hidden identity-based signatures also add attributes
to the credentials, with the aim of preventing the opener (the trusted party who
can deanonymize signers) from having to rely on the issuer in order to return
meaningful identifying information when opening a signature \cite{ks07}. Some AC
schemes also incorporate some sort of utility after authentication, like
AC schemes with revocation \cite{cks10}, or with ``watchlists'' \needcite.

@@

%% Main drawbacks of precedents
\paragraph{Limitations of GS and AC.} %
However, even the generalisation efforts that have independently taken place in
GS and AC have limited applicability in the real world. On the GS side, even
though MPS achieves a powerful balance on when a signature is
deanonymizable and what information can be extracted from it, upon closer look,
this information is somehow meaningless. Namely, the notion of identity is not
well defined and, thus, it is not possible to assess whether the value output by
their generalised open function makes sense or even if it is legitimate. Also,
just being able to prove membership at signing time is typically not enough. In
other words, in GS, utility at authentication time is very limited -- but this
is precisely the main strength in AC schemes, which also have a well defined
notion of identity via attributes. On the other hand, the notion of
accountability  is barely present in AC schemes. For instance, non-frameability
is not even considered, and it is non-frameability what prevent honest users
from being held accountable of dishonest actions they did not take part in.
%
In any case, perhaps the most impractical limitation of all is that too many
models exist in both fields. Although similar, each model focuses on slightly
different utility variants, making them hardly comparable. To the best of our
knowledge, no model exists that combines both fields, in a general enough
manner. In practice, needing to resort to too specific models forces system
designers to decide which feature they do not want to lose -- or, which one they
will lose.

@@

But the flexibility obtained from combining models from the AC and GS
domains into one also brings complexity. For instance, in order to meet
the general practice in AC, our UAS model includes multiple issuers. Along with
the combination of GS features, this paves the way to including multiple
openers, each with its own opening functionality. While powerful from a
practical point of view, this introduces similar (unavoidable?) limitations
privacy-wise as in functional encryption schemes \cite{bsw11} -- e.g., revealing
too granular information at signing or opening time reduces privacy. However, on
this occasion, signers controls ``how deanonymizable'' their signatures will
be, and know it at signature generation time. This flexibility on the opening
values also leads to the need of using extraction-based techniques for the
modelling, since we cannot expect that opening outputs the real identity of the
signer. This is no surprise, though, as it was already observed in GS schemes
with non-conventional opening \cite{dl21}, in DAA schemes \cite{cdl16,cdl16b}.
In fact, simulation-extractability seems to be most useful here, something that
was also observed in \cite{cl06} for producing signatures out of knowledge of
witnesses for concrete NP statements.

@@

\paragraph{Universality.}
To formally showcase that UAS can indeed claim to be universal, we prove that
restrictions of our generic construction imply the GS and AC ancestors of UAS.
But this includes not only their vanilla versions, but also relevant variants,
such as MPS \cite{ngsy22}, (a modified version of) group signatures with
message-dependent opening \cite{seh+12}, or anonymous credentials with
revocation \cite{cks10} or delegation \cite{bcc+09}.
%
Strictly speaking, for AC schemes, we build on a simple transform of our main
construction and model that turns non-interactive signatures into interactive
authentication protocols with protection against replays\footnote{This is
  the well-known folklore approach that has the verifier send a fresh nonce to
  be signed by the alleged owner of the private key related to a known
  public key.}. We further build ring signatures (RS) in the model given by
\cite{bkm06}, from an alternative UAS construction that improves issuer privacy.
%
The most interesting part being that, while we prove implications with
previously known schemes, our approach allows to merge their features: e.g.,
one can easily instantiate GSs with with no revocation at all, ACs
with full de-anonymization, RS with linkability, etc.

>>>>

%% Previous works summary w/ refs directly related contribs of UAS
%% (or, tech approaches of main precedents)
\paragraph{Evolution of AC and GS.} %
AC and GS systems have followed separate paths, with some exceptions.
Functionality-wise, Attribute-Based Group Signatures add attributes to the
credentials issued to group members, who can later use them to prove pre-fixed
claims \cite{emo09,aa14}. Hidden identity-based signatures also add attributes
to the credentials, with the aim of preventing the opener (the trusted party who
can deanonymize signers) from having to rely on the issuer in order to return
meaningful identifying information when opening a signature \cite{ks07}. Some AC
schemes also incorporate some sort of utility after authentication, like
AC schemes with revocation \cite{cks10}.

@@

%% Main drawbacks of precedents
\paragraph{Limitations of GS and AC.} %
However, even the generalisation efforts that have independently taken place in
GS and AC have limited applicability in the real world. On the GS side, even
though MPS achieves a powerful balance on when a signature is
deanonymizable and what information can be extracted from it, upon closer look,
this information is somehow meaningless. Namely, the notion of identity is not
well defined and, thus, it is not possible to assess whether the value output by
their generalised open function makes sense or even if it is legitimate. Also,
just being able to prove membership at signing time is typically not enough. In
other words, in GS, utility at authentication time is very limited -- but this
is precisely the main strength in AC schemes, which also have a well defined
notion of identity via attributes. On the other hand, the notion of
accountability  is barely present in AC schemes. For instance, non-frameability
is not even considered, and it is non-frameability what prevents honest users
from being held accountable of dishonest actions they did not take part in.
%
In any case, perhaps the most impractical limitation of all is that too many
models exist in both fields. Although similar, each model focuses on slightly
different utility variants, making them hardly comparable. To the best of our
knowledge, no model exists that combines both fields, in a general enough
manner. In practice, needing to resort to too specific models forces system
designers to decide which feature they do not want to lose -- or, which one they
will lose.

@@

But the flexibility obtained from combining models from the AC and GS
domains into one also brings complexity. For instance, in order to meet
the general practice in AC, our UAS model includes multiple issuers. Along with
the combination of GS features, this paves the way to including multiple
openers, each with its own opening functionality. While powerful from a
practical point of view, this introduces similar (unavoidable?) limitations
privacy-wise as in functional encryption schemes \cite{bsw11} -- e.g., revealing
too granular information at signing or opening time reduces privacy. However, on
this occasion, signers control ``how deanonymizable'' their signatures will
be, and know it at signature generation time. This flexibility on the opening
values also leads to the need of using extraction-based techniques for the
modelling, since we cannot expect that opening outputs the real identity of the
signer. This is no surprise, though, as it was already observed in GS schemes
with non-conventional opening \cite{dl21}, in DAA schemes \cite{cdl16,cdl16b}.
In fact, simulation-extractability seems to be most useful here, something that
was also observed in \cite{cl06} for producing signatures out of knowledge of
witnesses for concrete NP statements.

@@

\paragraph{Universality.}
To formally showcase that UAS can indeed claim to be universal, we prove that
restrictions of our generic construction imply the GS and AC ancestors of UAS.
This includes not only their vanilla versions, but also relevant variants,
such as MPS \cite{ngsy22}, (a modified version of) group signatures with
message-dependent opening \cite{seh+12}, or anonymous credentials with
revocation \cite{cks10} or delegation \cite{bcc+09}.
%
Strictly speaking, for AC schemes, we build on a simple transform of our main
construction and model that turns non-interactive signatures into interactive
authentication protocols with protection against replays\footnote{This is
  the well-known folklore approach that has the verifier send a fresh nonce to
  be signed by the alleged owner of the private key related to a known
  public key.}. We further build ring signatures (RS) in the model given by
\cite{bkm06}, from an alternative UAS construction that improves issuer privacy.
%
The most interesting part being that, while we prove implications with
previously known schemes, our approach allows to merge their features: e.g.,
one can easily instantiate GSs with with no revocation at all, ACs
with full de-anonymization, RS with linkability, etc.

### 2022-10-03 prism-backend/research/gsac/preliminaries-paper.tex
Pass.

\paragraph{Public-Key Encryption.} %
Defined as a tuple (\ESetup,\EKeyGen,\EEnc,\EDec). Algorithm $\parm \gets
\ESetup(\secpar)$ produces public parameters for the other algorithms.
$(\ek,\dk) \gets \EKeyGen(\parm)$ generates the encryption-decryption key
pair, algorithm $\Ec \gets \EEnc(\ek,\msg)$ encrypts message \msg with
encryption key \Eek, producing ciphertext \Ec, and $\msg/\bot \gets \EDec(\dk,
\Ec)$ decrypts ciphertexts using decryption key \Edk. We rely on IND-CCA
public-key encryption schemes.

@@

\paragraph{Signatures over Blocks of Committed Messages, with proofs.} %
We use signature schemes that allow signing blocks of messages, and commitments
to blocks of messages, and which are furthermore compatible with (efficient)
proof systems over the produced signature and signed (commitments to) messages.
For this purpose, we define such schemes as a tuple (\SBCMSetup,\SBCMKeyGen,
\SBCMBlind,\SBCMSign,\SBCMUnblind,\SBCMVerify).
%
Algorithm $\parm \gets \SBCMSetup(\secpar)$ produces the public parameters for
the scheme.
% 
$(\vk,\sk) \gets \SBCMKeyGen(\parm)$ produces a verification-signing key pair.
%
$(\Ccom,\pi,r) \gets \SBCMBlind^{\NIZKRel}(\vk,\osmsg,\smsg;x',w')$ is run by a
user who wants to request a signature over $\osmsg \cup \smsg$, where only \smsg
are revealed to the signer, and $\osmsg$ are to be signed in committed form.
Moreover, \osmsg (and, possibly, $w'$) are witnesses to some statement defined
over \smsg (and, possibly, $x'$), for a relationship \NIZKRel.
%
$\SBCMbsig \gets \SBCMSign^{\NIZKRel}(\sk,\Ccom,\pi,\smsg; x')$ is run by the
signer, to produce a partial signature \SBCMbsig over a set of committed
messages and set \smsg, satisfying relation \NIZKRel.
%
$\SBCMsig \gets \SBCMUnblind^{\NIZKRel}(\vk,\SBCMbsig,\Ccom,r,\osmsg,\smsg)$ is
run by the user to complete a partial signature \SBCMbsig received by the
signer, over \osmsg (committed to in \Ccom) and \smsg.
%
Finally, algorithm $1/0 \gets \SBCMVerify(\vk,\sig,\osmsg,\smsg)$ verifies a
signature \sig over message vector $\osmsg \cup \smsg$.
%
In addition, the produced signatures must be compatible with (efficient) NIZK
proofs of knowledge of a signature, and of (arbitrary) claims over the signed
(committed) messages. An \SBCM scheme must provide one-more-forgery type of
security.

@@

We follow common notation habits. $y \gets A(x)$ means that algorithm $A$, on
input $x$, produces output $y$. When $A$ is a probabilistic process, and we want
to make explicit the randomness $r$, we use $y \gets A(x;r)$. $\langle y_A,y_B
\rangle \gets \langle A(x_A),B(x_B)\rangle$ denotes algorithms $A$ and $B$,
running an interactive protocol, where $A$ receives $x_A$ as input and produces
$y_A$ as output, and $B$ receives $x_B$ and produces $y_B$.
%
Our generic constructions rely on the previously defined building blocks, which
share common (but different)
tokens: e.g., we resort to several NIZK proofs instances. To distinguish
them, we may use subscripts or superscripts: e.g., $\NIZKRel_A$ denotes the
NP relation used within some algorithm $A$, $\NIZKSetupRel$ means that
the setup algorithm of an underlying NIZK system is performed associated to
relation $\NIZKRel$; $\parm_X$ are the public parameters for scheme $X$, etc.
Variables written in bold font denote sets, e.g., while we use \ipk to denote a
single issuer public key, \sipk denotes a set of issuer's public keys --
although, when needed, we use explicit notation $\lbrace \ipk_i
\rbrace_{i\in[n]}$.
%
We also assume the existence of a well defined attribute space over which
issuers in our schemes can issue credentials on. We denote this set with
\AttrSpace.
%
$0(p)$ returns $True$ if $p=0$, or $False$ otherwise.

>>>>

\paragraph{Public-Key Encryption.} %
Defined as a tuple (\ESetup,\EKeyGen,\EEnc,\EDec). Algorithm $\parm \gets
\ESetup(\secpar)$ produces public parameters for the other algorithms.
$(\ek,\dk) \gets \EKeyGen(\parm)$ generates the encryption-decryption key
pair, algorithm $\Ec \gets \EEnc(\ek,\msg)$ encrypts message \msg with
encryption key \ek, producing ciphertext \Ec, and $\msg/\bot \gets \EDec(\dk,
\Ec)$ decrypts ciphertexts using decryption key \Edk. We rely on IND-CCA
public-key encryption schemes.

@@

\paragraph{Signatures over Blocks of Committed Messages, with proofs.} %
We use signature schemes that allow signing blocks of messages, and commitments
to blocks of messages, and which are furthermore compatible with (efficient)
proof systems over the produced signature and signed (commitments to) messages.
For this purpose, we define such schemes as a tuple (\SBCMSetup,\SBCMKeyGen,
\SBCMBlind,\SBCMSign,\SBCMUnblind,\SBCMVerify).
%
Algorithm $\parm \gets \SBCMSetup(\secpar)$ produces the public parameters for
the scheme.
% 
$(\vk,\sk) \gets \SBCMKeyGen(\parm)$ produces a verification-signing key pair.
%
$(\Ccom,\pi,r) \gets \SBCMBlind^{\NIZKRel}(\vk,\osmsg,\smsg;x',w')$ is run by a
user who wants to request a signature over $\osmsg \cup \smsg$, where only \smsg
are revealed to the signer, and $\osmsg$ are to be signed in committed form.
Moreover, \osmsg (and, possibly, $w'$) are witnesses to some statement defined
over \smsg (and, possibly, $x'$), for a relationship \NIZKRel.
%
$\bsig \gets \SBCMSign^{\NIZKRel}(\sk,\Ccom,\pi,\smsg; x')$ is run by the
signer, to produce a partial signature \bsig over a set of committed
messages and set \smsg, satisfying relation \NIZKRel.
%
$\sig \gets \SBCMUnblind^{\NIZKRel}(\vk,\bsig,\Ccom,r,\osmsg,\smsg)$ is
run by the user to complete a partial signature \bsig received by the
signer, over \osmsg (committed to in \Ccom) and \smsg.
%
Finally, algorithm $1/0 \gets \SBCMVerify(\vk,\sig,\osmsg,\smsg)$ verifies a
signature \sig over message vector $\osmsg \cup \smsg$.
%
In addition, the produced signatures must be compatible with (efficient) NIZK
proofs of knowledge of a signature, and of (arbitrary) claims over the signed
(committed) messages. An \SBCM scheme must provide one-more-forgery type of
security.

@@

We follow common notation habits. $y \gets A(x)$ means that algorithm $A$, on
input $x$, produces output $y$. When $A$ is a probabilistic process, and we want
to make explicit the randomness $r$, we use $y \gets A(x;r)$. $\langle y_A,y_B
\rangle \gets \langle A(x_A),B(x_B)\rangle$ denotes algorithms $A$ and $B$,
running an interactive protocol, where $A$ receives $x_A$ as input and produces
$y_A$ as output, and $B$ receives $x_B$ and produces $y_B$.
%
Our generic constructions rely on the previously defined building blocks, which
share common (but different)
tokens: e.g., we resort to several NIZK proofs instances. To distinguish
them, we may use subscripts or superscripts: e.g., $\NIZKRel_A$ denotes the
NP relation used within some algorithm $A$, $\NIZKSetupRel$ means that
the setup algorithm of an underlying NIZK system is performed associated to
relation $\NIZKRel$; $\parm_X$ are the public parameters for scheme $X$, etc.
Similarly, schemes algorithms are prefixed with acronyms identifying the scheme
(e.g., \ESetup, \NIZKSetup), which we may omit when clear from context.
Variables written in bold font denote sets, e.g., while we use \ipk to denote a
single issuer public key, \sipk denotes a set of issuer's public keys --
although, when needed, we use explicit notation $\lbrace \ipk_i
\rbrace_{i\in[n]}$.
%
We also assume the existence of a well defined attribute space over which
issuers in our schemes can issue credentials on. We denote this set with
\AttrSpace.
%
$0(p)$ returns $True$ if $p=0$, or $False$ otherwise.
